### 2014-03-25 paper/abstract.tex
start of video draft, figure images, pen holder files

\begin{abstract}
3D printers offer extraordinary flexibilty in prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for adding interactivity by removing interior material to form pipes and cavities. These can be used to sense input and generate haptic as well as visible output. We describe the design space of pipes and hollow chambers for interaction design: a variety of openings, topologies, and inserted media for pipes create diverse inputs and outputs.  We present a technique and design tool for routing pipes of various topologies through the interior of 3D printed parts.  Our design tool is integrated into a 3D model manipulation program.  We use two distinct routing algorithms.  One, for creating particular points of interaction or integrating with existing electronics, allows users to select begin and end points for pipes, then uses A* path routing and physics-based simulation to minimize the bending energy of routed paths.  The second, for creating, e.g., novel neon signs, allows users to enter a description of paths to follow: for this we developed a novel routing algorithm and offer proof of key traits.  We present several totally tubular prototypes we created using this tool to show its flexibility and potential, as well as to explore new points in the pipe design space.
\end{abstract}

>>>>

\begin{abstract}
3D printers offer extraordinary flexibilty in prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for adding interactivity by removing interior material to form pipes and cavities. These can be used to sense input and generate haptic as well as visible output. We describe the design space of pipes and hollow chambers for interaction design: a variety of openings, topologies, and inserted media for pipes create diverse inputs and outputs.  We present a technique and design tool for routing pipes of various topologies through the interior of 3D printed parts.  Our design tool is integrated into a 3D model manipulation program.  We use two distinct routing algorithms.  One, for creating particular points of interaction or integrating with existing electronics on the model's surface, allows users to select begin and end points for pipes, then uses A* path routing and physics-based simulation to minimize the bending energy of routed paths.  The second, for creating, e.g., novel neon signs, has users supply a desired internal shape which we fit a pipe route to: for this we developed a novel routing algorithm and offer proof of key traits.  We present several totally tubular prototypes we created using this tool to show its flexibility and potential, as well as to explore new points in the pipe design space.
\end{abstract}

### 2014-03-25 paper/designspace.tex
start of video draft, figure images, pen holder files

Pipes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' to have its contents physically accessible to an interacting user at interaction time.  This includes, for example, pipes which function as fountains, but it does not include pipes whose ends have been capped to keep water inside (whether this capping was done at print time, or later as a manual modification).  Pipes whose contents are not physically accessible to the user are ``closed on the user side''.  A pipe which is ``open on the system side'' is physically open to the machine, and one which is ``closed on the system side'' has its contents physically cut off from the machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents.

@@

Semi-closed pipes are open at the system end (for control of the enclosed medium) and closed at the user end.  The closed interface  may be fabricated from a mobile material: for example, a series of pipes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback, or a hinged cap can be blown open by air pressure.  Semi-closed pipes can also function as audio-generating resonance chambers.

@@

@@

``Gas'' comprises all compressible fluids.  Use of fluid pressure inside pipes can create haptic feedback at semi-closed interfaces, or the gases can be used as carriers for scents or fog.  As in \cite{Slyper-pressure}, structures can be engineered to change in air pressure when manipulated correctly (e.g., a spiral that changes pressure when twisted, but not pressed), and thus fluid pressure can also be used as an input.

>>>>

Pipes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  Since we are interested in interactive devices, we distinguish a ``user side" - part of the surface or interior of an object facing its user; and a ``system side" - part of the object (either interior or exterior) where other hardware such as electronics will be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' to have its contents physically accessible to an interacting user at interaction time.  This includes, for example, pipes which function as fountains, but it does not include pipes whose ends have been capped to keep water inside (whether this capping was done at print time, or later as a manual modification).  Pipes whose contents are not physically accessible to the user are ``closed on the user side''.  A pipe which is ``open on the system side'' is physically open to the machine, and one which is ``closed on the system side'' has its contents physically cut off from the machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents. \bjoern{wordy paragraph}

@@

Semi-closed pipes are open at the system end (for control of the enclosed medium) and closed at the user end.  The closed interface  may be fabricated from a malleable material: for example, a series of pipes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback, or a hinged cap can be blown open by air pressure.  Semi-closed pipes can also function as audio-generating resonance chambers.

@@

\bjoern{This section in general may need a small multipart figure that shows example photos or at least illustrations of these points in the design space. (Fig 2 is the abstract representation of the dimensions; here we have some concrete examples of points in the space).

Or maybe this comes later? it's a bit hard to follow these descriptions in the abstract.
}

@@

``Gas'' comprises all compressible fluids.  Use of fluid pressure inside pipes can create haptic feedback at semi-closed interfaces, or the gases can be used as carriers for scents or fog.  As in \cite{Slyper-pressure}, structures can be engineered to change in air pressure when manipulated correctly (e.g., a spiral that changes pressure when twisted, but not pressed), and thus fluid pressure can also be used as an input.  \bjoern{Isn't this the same example as two paragraph above?}

### 2014-03-25 paper/introduction.tex
start of video draft, figure images, pen holder files

The use of pipes can allow for \emph{redirection} and \emph{centralization} of active components, which creates \emph{iterable} devices.  For example, to create a touch-sensitive toy as in Figure \ref{fig:teaser}, a maker can select several locations of interest on the surface of the brain.  After creating pipes to those locations and printing her object, she can fill the pipes with conductive paint.  For sensing, she need only connect a single wire to a shared out of all the tubes, and can distinguish them via Swept Frequency Capacitve Sensing \cite{Sato-touche}.  If she decides that another area of interest is important to distinguish, she can simply print a new model with a new pipe leading to that location: the electronics assembly remains the same and she need only connect the same single wire to the tube out.  \valkyrie{not clear that this is our contribution.}

@@

The generation of these pipes, however, is not trivial.  Pipes need to be as straight as possible to ease the insertion of media post-print.  Independent pipes should not intersect, as this could lead to electrical shorting or other problems.  Pipes should not pass too close to the surface (unless they are being used for display purposes).  Consumer-grade 3D printers do not guarantee air- or water-tightness of their prints, so more than one layer may be required to prevent fluid leakage.  We have developed techniques for minimizing the bending energy of, enforcing non-intersection with, and avoiding the surface for pipes.  These techniques are embodied in our software tool to enable makers to create pipe-powered interfaces.

@@

\begin{itemize}
\item We lay out a design space of pipes and their types, contained media, topologies, and design focus.
\item We describe the interaction possibilities of pipes, including prior work as well as uncharted possibilities.
\item We offer algorithms and techniques for routing pipes, as well as a design tool implementing them.
\item We showcase a set of examples, enabled by our modeling tool and exploring new points in the design space.
\end{itemize}

>>>>

The use of pipes can allow for \emph{redirection} and \emph{centralization} of active components to various locations on a printed object's surface.   For example, to create a touch-sensitive toy as in Figure \ref{fig:teaser}, a maker can select several locations of interest on the brain.  After creating pipes to those locations and printing her object, she can fill the pipes with conductive paint.  For sensing, she need only connect a single wire to a shared out of all the tubes, and can distinguish them via Swept Frequency Capacitve Sensing \cite{Sato-touche}.  In this way, she has only a single active element (the microcontroller setup), but her otherwise-inert 3D printed object (the brain) has several active sensing locations.  This can also be used for rapid iteration of designs.  Our maker need only add a digital tube, reprint her brain, and re-fill it with copper paint to have a new version; no complicated electromechanical assembly required.

@@

The generation of these pipes, however, is not trivial.  Pipes need to have maximum possible bend radius to ease the insertion of media post-print.  Independent pipes should not intersect, as this could lead to electrical shorting or other problems.  Pipes should not pass too close to the surface (unless they are being used for display purposes).  Consumer-grade 3D printers do not guarantee air- or water-tightness of their prints, so more than one layer may be required to prevent fluid leakage.  Pipes for applications like neon signs must have an Euler tour through them: this allows a single EL wire to pass through every edge on the pipe graph.  We have developed techniques for minimizing the bending energy of, enforcing non-intersection with, and avoiding the surface for pipes, as well as expanding user-specified paths and generating Euler tours.  These techniques are embodied in our software tool to enable makers to create pipe-powered interfaces.

@@

\begin{itemize}
\item We lay out a design space of pipes and their openings, contained media, topologies, and design focus.
\item We offer algorithms and techniques for routing pipes, as well as a design tool implementing them.
\item We describe the interaction possibilities of pipes, including prior work as well as uncharted possibilities.
\item We showcase a set of examples, enabled by our modeling tool and exploring new points in the design space.
\end{itemize}

### 2014-03-25 paper/relatedwork.tex
start of video draft, figure images, pen holder files

While using a series of tubes can replicate a variety of interaction techniques, our work's \emph{contribution} is more generally related to two existing research areas: fabrication and routing.

@@

Previous work has investigated the integration of interactivity with the fabrication process.  \cite{Willis-printedoptics} used light pipes, solid clear tubes, to create integrated displays: our work allows a superset of these capabilities, and Printed Optics is a part of our design space.  Willis, et al., also did not offer a design tool or techniques for creating light pipes automatically.  \cite{Savage-sauron} investigated the use of computer vision to track interactions with physical mechanisms.  Using tubes for redirection, it is possible to replicate these mechanically sensitive techniques (see Figure \ref{fig:pens}), however our work does not focus on computer vision alone.  Slyper, et al., \cite{Slyper-pressure} 3D printed flexible robot armatures capable of sensing interaction via changes in air pressure.  This technique is complementary to ours, as we allow for placing these interactive elements anywhere on an object's surface while they had to place them in locations accessible for electronics insertion.

@@

Other research has attempted to bring traditional electronics to 3D prints, either on the surface or via interior channels.  Sells, et al., in \cite{Sells-reprap} create channels on the surface of a 3D printed object which are filled with a special computer-controlled syringe of liquid solder to create connections.  Their focus is on creating self-replicating 3D printers, and their electronics are combined to traditional 2D routed circuitboards.  Sarik, et al., \cite{Sarik-tracebrush} \valkyrie{I never found a real paper of this.  The Microsoft website has a "print" of it with TEI boilerplate, but I couldn't find it in the digital library.  What's up with that?} and the Optomec company \cite{optomec} both created techniques for taking an already-printed object and spraying conductive material onto its surface.  In contrast, our technique can lead to more durable circuitry, since it is routed through the core of the object rather than on its surface.  Navarette, et al., in \cite{Navarrette-gps} created a functioning GPS whose circuit routing extended into true 3D (as compared to the 2.5D boards created by stacking traditional 2D circuitboards).  However, they created this one-off routing by hand and do not offer any general algorithms or techniques.  In the area of robotics, softer-than-skin interactive surfaces have been created using tubed interfaces fabricated from silicone molds.  \cite{Park-microchannels} and \cite{Majidi-curvature} use microchannels filled with liquid metal for pressure or curvature sensing: the metal in the channels changes resistance as the channels are stretched.  These techniques have not been used with fabrication, nor has a design tool to create such sensors been previously described.

@@

General-purpose routing for connecting 2D capacitive sensors to terminals is discussed in \cite{Savage-midas}.  These techniques, based on circuitboard routing, are limited to planar and unfoldable shapes, and are intended for use on the surface of objects rather than through their cores.

@@

Prior work on modeling the interior of objects (as opposed to their exterior) has been limited.  \valkyrie{ryan, can you give citations here?}  Sauron automatically modified the interior of existing mechanical objects to make them amenable to vision-based sensing \cite{Savage-sauron}, while Willis, et al.,'s Infrastructs project created encoded interior cavities in 3D printed objects designed to be sensed with a terahertz imaging tool.  Our work differs in that we want to allow users to design pipes and integrate interactive objects with them.

>>>>

Our contributions relate to other efforts to add interactivity to digitally fabricated objects; and to techniques for routing and mode modification.  \bjoern{The related work section is taking shape but still has a bit of a "missing the forest for the trees" feeling.)}

\bjoern{Tactically, to communicate who would be is a qualified reviewer for this paper in the UIST community, we should make sure to have appropriate references to Baudisch, Poupyrev, Harrison in here}

@@

Previous work has investigated the integration of interactivity with the fabrication process.  Printed Optics (\cite{Willis-printedoptics}) used light pipes, solid clear tubes, to create integrated displays: our work is inspired by this and goes towards a design tool for these capabilities \valkyrie{bad phrasing; don't de-fang the tool at the front of the paper!}; Printed Optics is a part of our design space, which includes other uses of pipes.  Willis, et al., also did not offer a design tool or techniques for creating light pipes automatically.  Sauron (\cite{Savage-sauron}) investigated the use of computer vision to track interactions with physical mechanisms.  While Savage, et al., used mirrors for redirection, we believe that the required line-of-sight operation for their computer vision techniques may be possible using printed optics or inserted optical fibers.  Slyper, et al., \cite{Slyper-pressure} 3D printed flexible robot armatures capable of sensing interaction via changes in air pressure.  This technique is complementary to ours, as we allow for placing these interactive elements anywhere on an object's surface while they had to place them in locations accessible for electronics insertion.

@@

Other research has attempted to bring traditional electronics to 3D prints, either on the surface or via interior channels.  Sells, et al., in \cite{Sells-reprap} create channels on the surface of a 3D printed object which are filled with a special computer-controlled syringe of liquid solder to create connections.  Their focus is on creating self-replicating 3D printers, and their electronics are confined to traditional 2D routed circuitboards.  Sarik, et al., \cite{Sarik-tracebrush} \valkyrie{I never found a real paper of this.  The Microsoft website has a "print" of it with TEI boilerplate, but I couldn't find it in the digital library.  What's up with that?} and the Optomec company \cite{optomec} both created techniques for taking an already-printed object and spraying conductive material onto its surface.  In contrast, our technique can lead to more durable circuitry, since it is routed through the core of the object rather than on its surface.  Navarette, et al., in \cite{Navarrette-gps} created a functioning GPS whose circuit routing extended into true 3D (as compared to the 2.5D boards created by stacking traditional 2D circuitboards).  However, they created this one-off routing by hand and do not offer any general algorithms or techniques.  In the area of robotics, softer-than-skin interactive surfaces have been created using tubed interfaces fabricated from silicone molds.  \cite{Park-microchannels} and \cite{Majidi-curvature} use microchannels filled with liquid metal for pressure or curvature sensing: the metal in the channels changes resistance as the channels are stretched.  These techniques have not been used with fabrication, nor has a design tool to create such sensors been previously described.

@@

General-purpose routing for connecting 2D capacitive sensors to terminals is discussed in Midas (\cite{Savage-midas}) \bjoern{Maybe move discussion of Midas to prior section since it really is about fabricating interactive devices. And then here just quickly review that lots of work has gone into routing algorithms for PCBs but that they assume multiple stacked planes connected with vertical vias, not general 3d routing.}.  These techniques, based on circuitboard routing, are limited to planar and unfoldable shapes, and are intended for use on the surface of objects rather than through their cores.

@@

Prior work on modeling the interior of objects (as opposed to their exterior) has been limited.  \valkyrie{ryan, can you give citations here?  I know there is that one about balancing, but Tovi mentioned a ``fruit slicing" one?}  In Make it Stand, material is removed from the interior of objects with the goal of changing their center of gravity \cite{Prevost-makeitstand}.  Bickel, et al., explored enforcing deformation behavior with material blending \cite{Bickel-deformation}, while B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}.  Sauron automatically modified the interior of existing mechanical objects to make them amenable to vision-based sensing \cite{Savage-sauron}, and Infrastructs created encoded interior cavities in 3D printed objects designed to be sensed with a terahertz imaging tool.  Our work differs in that we want to allow users to design pipes and integrate interactive objects with them.
